### 2020-03-25 Paper_shortened_pdf.Rmd
Editor sex comment

Child sex differences are relevant in the interplay between parenting and emotional adjustment. Parents tend to raise boys and girls differently; expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999]. Furthermore, social learning theory posits that the same-sex parent plays a stronger role in socialization; particularly for sex-typed behavior, such as emotion display rules [@busseySocialCognitiveTheory1999]. Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a]. We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning towards girls than boys, and more anger towards boys than girls, $H_{7a-d}$. Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent, $H_{8a-d}$. Sex differences in emotional adjustment have been found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a]. However, a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a]. Consequently, we have no hypothesis regarding sex differences in average levels of emotional adjustment.

@@

We found sex differences in the random intercepts of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)). Counter to hypotheses $H_{7b,c}$, there were no sex differences in consistency, and both parents displayed more inductive reasoning towards sons. We found no sex differences in the random intercept of emotional adjustment. Hypotheses $H_{8a-d}$ were rejected: There was no evidence for stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent. The only predictive effects moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning. Emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls. In other words, if children display worse emotional adjustment than usual in one wave, parents increase inductive reasoning more strongly in the next wave for boys than for girls.

@@

In line with prior research, we found that parents raise boys and girls differently. Mothers and fathers both displayed more anger and inductive reasoning with sons than with daughters, and fathers displayed less warmth towards sons. These findings suggest that parents are stricter towards their sons. Parenting effects were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors. Child effects were moderated by child sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male. When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.

>>>>

Child sex differences are considered to be an important covariate in the literature on parenting and emotional adjustment. Parents raise boys and girls differently, expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999]. Furthermore, the same-sex parent is thought to play a stronger role in socialization [@busseySocialCognitiveTheory1999]. Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a]. We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning, and less anger, towards girls than boys, $H_{7a-d}$. We also predict associations between qualitative parenting and emotional adjustment to be stronger between children and their same-sex parent, $H_{8a-d}$. Finally, although sex differences in emotional adjustment have been found in older children [@vanlissaRoleFathersMothers2018a], there is little evidence for such differences in young children [@else-questGenderDifferencesTemperament2006a].

@@

We found sex differences in the average levels of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)). Counter to hypotheses $H_{7b,c}$, there were no sex differences in consistency, and both parents displayed more inductive reasoning towards sons. We found no sex differences in the average level of emotional adjustment. Hypotheses $H_{8a-d}$ were rejected: There was no evidence for stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent. The only predictive effects moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning: When children displayed worse emotional adjustment than usual in one wave, their parents reported stronger increases in inductive reasoning in the next wave for boys than for girls.

@@

In line with prior research, we found that both parents displayed more anger and inductive reasoning, and less warmth, towards sons than towards daughters. These findings suggest that parents are stricter with sons. Sex did not moderate parenting effects, which suggests that boys and girls are similarly affected by qualitative parenting behaviors. Sex did moderate child effects, however: Emotional adjustment predicted both parents' inductive reasoning more negatively when children were male. This finding suggests that some of parents' greater strictness toward boys is alleviated when boys are well-adjusted.
